---
title: "Comparison between Horwitz-Thompson and H\'ajek Estimators"
date: 11.09.2018
output: html_document
---

```{r, echo = FALSE}
###### ------------------------------------------------------------
###### ------------------------------------------------------------
###### NAME: Horwitz-Thompson and Hajek estimators
###### DATE: September 2018
###### Version: 1
###### ------------------------------------------------------------
###### ------------------------------------------------------------
###### Script to understand the HT and H estimators to include them in the school simulation
###### Following: https://cran.r-project.org/web/packages/sampling/vignettes/HT_Hajek_estimators.pdf
if(!require(sampling)) install.packages("sampling",repos = "http://cran.us.r-project.org")
library("sampling")
```
# The Horwitz-Thompson and Haj\'ek Estimators

The Horwitz-Thompson estimator estimates the sum of a variable $y_i$ over a subgroup $s$ of a population $U$, weighing each item according to the probability $\pi_i$ to draw from the sample. 

$$\hat{y}_{HT}=\sum_{i\in s}{\frac{y_i}{\pi_i}}$$


# Setting up the simulation
```{r}
data("belgianmunicipalities")
attach(belgianmunicipalities)

# Define sample size
n <- 20
# Define inclusion probabilities
# Tot04 contains the Belgian population on the 1st of July of 2004
# The probability of inclusion is directly related to the number of inhabitants for each municipalities
# sampling::inclusionprobabilities defines a probability based on the values given by 'a', 
# proportional to the size of the sampling design given by 'n'
pik <- sampling::inclusionprobabilities(a = Tot04, n = n)
N <- length(pik)
# Now we look at the number of simulations
sim <- 10
s_ht <- s_h <- array(0, c(sim, 4))
```

## Case 1. Constant
The $y_i$ is constant for $i = 1, \cdots, N$.
```{r}
y1 <- rep(3, N)
```


## Case 2. Poisson
We use Poisson sampling to draw a sample $s$. 

```{r}
# TaxableIncome includes the total taxable income by municipality at 2004
y2 <- TaxableIncome
```


## Case 3. Model 1
The $y_i$ are generated using the following model: 
$$x_i = i,\  \pi_i = \frac{n \cdot x_i }{\displaystyle{\sum_{i=1}^N{x_i}}},\ y_i = \frac{1}{\pi_i}$$
In this case, the $\pi_i$ are weakly or negatively correlated to the $y_i$.  

```{r}
x <- 1:N
pik_3 <- inclusionprobabilities(x, n)
y3 <- 1/pik_3
```


## Case 4. Model 2
The $y_k$ are generated using the following model: 
$$x_i = i,\ y_i = 5(x_i + i),\ i \sim N(0, \frac{1}{3})$$
in this case the Horvitz-Thompson estimator should perform better than the H\'ajek
estimator.

```{r}
epsilon <- rnorm(N, 0, sqrt(1/3))
pik_4 <- pik_3
y4 <- 5*(x + epsilon)
```

# Simulation 

```{r,  echo = TRUE, results = "hide"}
# Define a vector to include the HT and the H estimators
ht <- numeric(4)
hajek <- numeric(4)

# Run the simulation for the predefined number of simulations before
i <- 1
for(i in 1:sim){
  # Indicate where we are in the simulation
  cat("Simulation ",i,"\n")
  # Case 1
  cat("Case 1\n")
  # samping::UPtille samples from a distribution of unequal probabilities, without replacement, with a fixed sample size
  # It uses the probabilities of inclusion defined above
  s <- sampling::UPtille(pik)
  # Calculating the Horwitz-Thompson estimator based on the sampling 's'
  ht[1] <- HTestimator(y1[s == 1], pik[s == 1])
  # Calculating the H??jek estimator based on the sampling 's'
  # 'N' is the size of the population to calculate the 'total' estimator, not the mean estimator
  hajek[1] <- Hajekestimator(y1[s == 1], pik[s == 1], N, type="total")
  
  # Case 2
  cat("Case 2\n")
  # Sampling using a poisson process
  s1 <- sampling::UPpoisson(pik)
  # Calculating the Horwitz-Thompson estimator based on the sampling 's1'
  ht[2] <- sampling::HTestimator(y2[s1 == 1], pik[s1 == 1])
  # Calculating the H??jek estimator based on the sampling 's1'
  hajek[2] <- sampling::Hajekestimator(y2[s1 == 1], pik[s1 == 1], N, type = "total")
  
  # Case 3
  cat("Case 3\n")
  # Calculating the Horwitz-Thompson estimator based on the sampling 's'
  ht[3] <-  sampling::HTestimator(y3[s == 1], pik_3[s == 1])
  # Calculating the H??jek estimator based on the sampling 's'
  hajek[3] <- sampling::Hajekestimator(y3[s == 1], pik_3[s == 1], N, type = "total")
  
  # Case 4
  cat("Case 4\n")
  # Calculating the Horwitz-Thompson estimator based on the sampling 's'
  ht[4] <- sampling::HTestimator(y4[s == 1], pik_4[s==1])
  # Calculating the H??jek estimator based on the sampling 's'
  hajek[4] <- sampling::Hajekestimator(y4[s == 1], pik_4[s == 1], N, type = "total")
  
  # Store the results in the simulation vectors
  s_ht[i,] <- ht
  s_h[i,] <- hajek
  }
```


